---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Scraping BM-EB

```{r scrape-bmeb}

# Install packages if not already installed
# install.packages("rvest")
# install.packages("xml2")

library(rvest)
library(dplyr)

# 1. Define the URL
url <- "https://data.bishopmuseum.org/ethnobotanydb/ethnobotany.php?b=list&o=1"

# 2. Read the HTML content
page <- read_html(url)

# 3. Inspect and extract all <table> elements
tables <- page %>% html_nodes("table")

# 4. Review how many tables are present
length(tables)
# print(head(html_table(tables[[1]], fill = TRUE)))

# 5. Assuming the table of interest is the first one:
ethno_table <- tables[[1]] %>% html_table(fill = TRUE)

# 6. View or process the resulting data frame
glimpse(ethno_table)




```




# ethnobotany-resources

<!-- badges: start -->
<!-- badges: end -->

The goal of ethnobotany-resources is to collect datasets, field notes, and literature references for plants used in traditional medicine and cultural
practices.

# Ethnobotanical Resources

This repository collects datasets, field notes, and literature references  for plants used in traditional medicine and cultural practices.

```{r echo=FALSE, message=FALSE}

library(knitr)

# Load the data from your repo
databases <- read.csv("data/ethnobotanical_databases.csv")

# Render as a black-and-white Markdown table
kable(databases, format = "markdown", align = "l")
knitr::opts_knit$set(root.dir = "C:/Users/Stacy/OneDrive/Desktop/ethnobotany-resources2")

```
